\setlength{\epigraphrule}{0\p@}
\setlength{\epigraphwidth}{.28\textwidth}
\epigraph{\large{\textit{To Carol and Rub\'{e}n.}}}{}

First of all, I want to thank my supervisor Isaac Salazar-Ciudad for guiding me in this scientific initiation journey. I especially appreciate all the discussions on evodevo papers (that made me more critical) and for encouraging me to search for my own solutions. 
\vspace{1ex}

I am grateful to the members of my thesis committee: Jukka Jernvall, Osamu Shimmi and Mikael Fortelius. They not only improved the quality of this work with their critical comments and suggestions, but they always treated me in a respectful way that made me feel valuable.
I am grateful to Academy of Finland for funding and to the Doctoral School Health Sciences (DSHealth) for awarding me with a doctoral dissertation grant.
I also want to show my gratitude to all the University and Biotechnology Institute Staff that helped and supported me in any way during these four years.
\vspace{1ex}

I was lucky to have Roland Zimm, Miquel Mar\'{i}n-Riera and Miguel Brun-Usan as colleagues through the entire PhD process. I especially want to thank Roland for his friendship, for all the good moments we shared and for the support he gave me, together with wise advices as "ein Bier ist kein Bier". I am going to miss all those geeky conversations.
%
I am indebted to all the people who were or still are in Salazar-Ciudad's group: Alexis, Pascal, Fernando, Lisandro, Jhon and Anton for their unconditional help and for creating a great working environment. I am specially grateful for the friendship forged with Alexis Matamoro-Vidal.
\vspace{1ex}

It was an honour for me to be part of the Helsinki EvoDevo community. It is always inspiring to see how Jukka loves science and how he can always see the things from different perspectives. I owe my deepest gratitude to Susanna Sova, Jacqueline Moustakas-Verho and Mona Christensen for their constant support and for the many laughs shared. I would also like to show my gratitude to the other members of the community. Thank you Tuomas A., Tuomas K., Elodie, Outi, Ritva, Yoland, Rishi, Teemu, Ian, Johann, and Mia.  

\vspace{1ex}
This thesis would not have been possible without the collaboration of Antonio Barbadilla, Marta Coronado-Zamora and David Castellano from the Universitat Aut\`{o}noma de Barcelona (UAB) and their expertise in bioinformatics and populations genetics. It was great to collaborate with them, their enthusiasm made everything easier.

\vspace{1ex}
I am grateful to all the people, inside and outside the university, that have shared part of this experience with me. It was great to interact with the Developmental Biology people. The Tv\"{a}rminne and Hyyti\"{a}l\"{a} meetings were always interesting (and fun). Taking the risk of forgeting someone, I want to thank Anne, Filipe, Julia, Fabien, Mar\'{i}a, Ewelina, Darshan, Alex, all friends in Barcelona, Carlos, Manuel and the Bio-ppl group.

\vspace{1ex}

I want to express my warmest gratitude to my parents and sisters that, despite the distance, have always been there for me.
Finally, I want to thank Carol for supporting me all the way and for bearing a long distance relationship for years. It was not easy, but all the effort has been worth it.