
\vspace{1cm}

This thesis is based on the following original publications which are referred to by their roman numerals in the text:

\vspace{0.5cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\begin{flushleft}
\leftskip3em
\rightskip\leftskip
\normalsize{

\begin{enumerate}[label=\Roman*]
  \item  \textbf{Salvador-Mart\'{i}nez, I.}, Salazar-Ciudad, I., 2015. 
  How complexity increases in development: An analysis of the spatial-temporal dynamics of 1218 genes in Drosophila melanogaster.
   \textit{Dev. Biol.} \textbf{405}, 328-339. doi:10.1016/j.ydbio.2015.07.003
  \vspace{0.5cm}
    
  \item  \textbf{Salvador-Mart\'{i}nez, I.}, Salazar-Ciudad, I. 
  How complexity increases in development: An Analysis of the Spatial-Temporal Dynamics in \textit{Ciona intestinalis}. Manuscript.
  \vspace{0.5cm}

  \item \textbf{Salvador-Mart\'{i}nez, I.\footnotemark[1]}, Coronado-Zamora, M.\footnotemark[1], Castellano, D.\footnotemark[1], Barbadilla, A., and Salazar-Ciudad, I. 
  Mapping Natural Selection within \textit{Drosophila melanogaster} embryo's anatomy.
   Manuscript.
  \vspace{0.5cm}
  
  \item  Coronado-Zamora, M.\footnotemark[1], \textbf{Salvador-Mart\'{i}nez, I.\footnotemark[1]}, Castellano, D., Barbadilla, A., and Salazar-Ciudad, I. 
  When in development: Estimating natural selection from the DGRP Estimating natural selection through \textit{Drosophila} development from population genomics
Manuscript.
  
\end{enumerate}
%
\footnotetext[1]{These authors contributed equally to this work.}
\renewcommand*{\thefootnote}{\arabic{footnote}}
%
}
\end{flushleft}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \vspace{2cm}
\subsubsection{Author contributions}
\small{
In article I, I. S-C. and I. S-M. conceived the study, I. S-M. analyzed the data and performed the statistical analysis, I. S-C. and I. S-M. wrote the manuscript. 

In article II, I. S-C and I. S-M. conceived and designed the study. I. S-M. analyzed the data and performed the statistical analysis, I. S-C. and I. S-M. wrote the manuscript. 

In article III, ISM analyzed gene age, gene expression and the relationship between codon usage gene expression and adaptation. MCZ calculated the selection estimators. DC help MCZ. ISC, AB , ISM and DC designed the study. ISC conceived the study. ISC wrote the article with the help of all co-authors.

In article IV, ISM analyzed the gene expression data through the life cycle and performed the clustering. ISM and MCZ performed the statistical analysis. MCZ calculated the selection estimators and the genomic determinants. DC help MCZ. ISC, AB, ISM and DC designed the study. ISC conceived the study. ISC wrote the article with the help of all co-authors. 
}